The \dealii Library, Version 9.4

### 2022-06-23 9.4/paper.tex
Remove duplicate p4est publication. Adjust p4est references.

\documentclass{ansarticle-preprint}
%\usepackage{ucs}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
%\usepackage{cite}
\usepackage{anslistings}
\usepackage{multicol}
\usepackage{pdfsync}

@@

\title{The \dealii{} Library, Version 9.4}

@@

   \author[4]{Marco~Feder}
\affil[4]{SISSA,
   International School for Advanced Studies,
   Via Bonomea 265,
   34136, Trieste, Italy.
   {\texttt{ \{marco.feder,luca.heltai\}@sissa.it}}}

@@

\author[13]{Rene~Gassm{\"o}ller}
\affil[13]{Department of Geological Sciences,
   University of Florida,
   1843 Stadium Road,
   Gainesville, FL, 32611, USA.
  {\texttt{rene.gassmoeller@ufl.edu}}}

\author[5]{Timo~Heister}
 \affil[5]{School of Mathematical and Statistical Sciences,
   Clemson University,
   Clemson, SC, 29634, USA
   {\texttt{heister@clemson.edu}}}

@@

 \author[6,7]{Martin~Kronbichler}
 \affil[6]{Department of Information Technology,
   Uppsala University,
   Box 337, 751\,05 Uppsala, Sweden.
   {\texttt{martin.kronbichler/simon.sticko@it.uu.se}}}
 \affil[7]{Institute of Mathematics,
   University of Augsburg,
   Universit\"atsstr.~12a, 86159 Augsburg, Germany.
   {\texttt{martin.kronbichler@uni-a.de}}}

\author[8]{Matthias~Maier}
\affil[8]{Department of Mathematics,
  Texas A\&M University,
  3368 TAMU,
  College Station, TX 77845, USA.
  {\texttt{maier@math.tamu.edu}}}

\author[7,9]{Peter Munch}
 \affil[9]{Institute of Material Systems Modeling,
 Helmholtz-Zentrum Hereon,
 Max-Planck-Str. 1, 21502 Geesthacht, Germany.
   {\texttt{peter.muench@hereon.de}}}

\author[10]{Jean-Paul~Pelteret}
\affil[10]{Independent researcher.
{\texttt{jppelteret@gmail.com}}}

\author[6,11]{Simon~Sticko}
\affil[11]{Department of Mathematics and Mathematical Statistics,
   Umeå University,
   SE-90187 Umeå, Sweden}

@@

\author[12]{David Wells}
\affil[12]{Department of Mathematics, University of North Carolina,
  Chapel Hill, NC 27516, USA.
  {\texttt{drwells@email.unc.edu}}}

@@

\dealii{} version 9.4.0 was released June 10, 2022.
\todo{Update release date.}
This paper provides an
overview of the new features of this release and serves as a citable
reference for the \dealii{} software library version 9.4. \dealii{} is an
object-oriented finite element library used around the world in the
development of finite element solvers. It is available for free under the
GNU Lesser General Public License (LGPL). Downloads are available at
\url{https://www.dealii.org/} and \url{https://github.com/dealii/dealii}.

The major changes of this release are:
%
\begin{itemize}
  \item Advances in simplex- and mixed-mesh support (see Section~\ref{sec:simplex});
  \item Repartitioning of distributed meshes (see Section~\ref{sec:repartitioning});
  \item Advances in matrix-free infrastructure (see Section~\ref{sec:mf});
  \item Advances in multigrid infrastructure (see Section~\ref{sec:multigrid});
  \item CutFEM support (see Section~\ref{sec:cut});
  \item Integration with the Computational Geometry Algorithms Library (CGAL) (see Section~\ref{sec:cgalwrappers});
  \item Performance improvements in the particle infrastructure (see Section~\ref{sec:particles});
  \item Two new tutorial programs and one new code gallery program (see
    Section~\ref{subsec:steps}).
\end{itemize}
%

While all of these major changes are discussed in detail in
Section~\ref{sec:major}, there
are a number of other noteworthy changes in the current \dealii{} release,
which we briefly outline in the remainder of this section:
%
\begin{itemize}
%  \item \texttt{AffineConstraints::make\_consistent\_in\_parallel()} allows
%  to make constraints consistent in parallel.
  \item The \texttt{DataOutResample} class interpolates values defined on one
  triangulation onto a second potentially unrelated triangulation.
  By using this class, one can output the result obtained on an
  unstructured mesh on a structured one (which might facilitate a more
  memory-efficient storage format, for example if this second
  triangulation is a uniformly refined rectangle or box), or one can create a slice in 3D.
  \item The new member function \texttt{find\_point\_owner\_rank()} of
        \texttt{parallel\allowbreak ::distributed::\allowbreak Tri\allowbreak angulation} allows one to find the MPI
  ranks of cells containing specified points.
  It is communication-free and leverages the functionality of p4est (>v.2.2).
  Its algorithm is described in \cite{burstedde2020parallel}. This information will
  enable efficient construction of the
  communication pattern used in the class \texttt{Utilties::\allowbreak MPI::\allowbreak RemotePointEvaluation}. Furthermore, this function could be used in the future to allow
  particle simulations in which particle movement is not
  limited by CFL conditions, as done in \cite{mirzadeh2016parallel}.
  \item The new function
  \texttt{GridGenerator::pipe\_junction()}
  generates a triangulation of three cone-shaped pipes that cross at a bifurcation point in any possible configuration.
  A manifold description is applied to the boundary, which can be extended into the volume via transfinite interpolation \cite{Gordon82} using the \texttt{TransfiniteInterpolationManifold} class \cite{dealII90}.
  \item A new DoF renumbering function \texttt{DoFRenumbering::support\_point\_wise()} which groups together
  shape functions by their support point. This functionality is useful in both developing nodal schemes since, e.g., the $x$, $y$, and $z$
  velocities at a point will be consecutive in the solution vector. It also improves interoperability with external libraries which expect
  data in this format.
  \item A new module \texttt{Utilities::MPI::LargeCount} which enables sending and receiving MPI messages containing more than $2^{31}$ objects.
  This library either uses the new MPI-4 functions, such as \texttt{MPI\_Send\_c()}, or an internal implementation of large objects for
  compatibility with MPI-3.
  \item The \texttt{FEInterfaceValues} class, which computes common quantities at the interface of two cells, has been overhauled to make it
  more consistent with the rest of the library and use more intuitive names for functions. For example,
  \texttt{FEInterfaceValues::jump\_gradient()} is now \texttt{FEInterfaceValues::jump\_in\_shape\_gradients()}. Several new
  functions, such as \texttt{FEInterfaceValues::get\_jump\_in\_function\_values()}, have also been added.
  \item Vectors attached to \texttt{DataOut} do not need to be in ghosted state anymore. Internally, we create a copy of the vector with appropriate ghosting.
\end{itemize}
%
The changelog lists more than 100 other features and bugfixes.

@@

This release of \dealii{} contains a number of large and significant changes,
which will be discussed in this section.
It of course also includes a
vast number of smaller changes and added functionality; the details of these
can be found
\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_3_0_and_9_4_0.html}
{in the file that lists all changes for this release}; see \cite{changes94}.

@@

\dealii{} has supported simplex and mixed meshes since the previous
release, 9.3, on an experimental basis.
We have continued to work on this support; the current state is
sufficient to support several larger-scale programs, although not all
functionality in \dealii{} works with such meshes yet. Specifically,
we have fixed many bugs and generalized existing functions that
previously only worked for hypercube-shaped cells. The most notable
new functionalities are:

@@

\dealii{} has two classes that support distributed storage of meshes: 
\texttt{parallel::\allowbreak distributed::\allowbreak Triangulation}
(which in the following we will abbreviate as \texttt{p::d::T}) and
\texttt{parallel::\allowbreak fully\allowbreak distributed::\allowbreak
  Triangulation} (or, in short, \texttt{p::f::T}). The former is
partitioned based on the space-filling Morton (or Z-order) curve
implemented by the \texttt{p4est} backend. The latter, more recent
class is currently statically partitioned at creation time.

@@

The matrix-free infrastructure of \dealii{} is used to solve problems
without assembling matrices, providing only the action of the matrix
instead. It has seen a number of new features in this release, the most
notable ones being:
\begin{itemize}
\item Improved support for computations with Hessians of shape functions: just like values and gradients, Hessians can be
now evaluated and integrated during matrix-free loops
both for cells and faces. This could enable, for example, writing
a matrix-free version of step-47, which solves the biharmonic equation with the discontinuous Galerkin method.
\item Cell-centric loops now also allow access to gradients and Hessians
of neighboring cells on faces. The major difficulty here was the relative orientation
of cells on unstructured meshes.
\item Users can now create their own cell batches, by providing \texttt{FEEvaluation::reinit()} a list of cell IDs. \texttt{FEEvaluation}
accesses the appropriate data and reshuffles mapping data accordingly on
the fly in order to enable vectorization over cells. The new feature is useful in several
contexts. Examples are simulations with sharp interfaces (e.g., two-phase flow
or shock capturing), where one needs to treat cells that are ``cut'' by
the interface in a special way. A challenge is that cell batches
might contain cut or not-cut cells, making vectorization of these operations potentially more complicated. Previous functionality has provided the option of masking certain cells in cell batches, which works well if
the code paths do not diverge too much. Another way is to
categorize cells during \texttt{MatrixFree::reinit()} in such a way that mixed cell batches do not occur. However, \texttt{MatrixFree::reinit()}
might be too expensive if recategorization needs to happen very frequently to follow the dynamics of a system, e.g., in each time step. Despite some overhead compared to static matrix-free loops, the new feature can be the best options in these scenarios.
\item Initial support for H(div)-conforming elements with Piola transform, based on Raviart--Thomas finite element spaces with the updated class \texttt{FE\_RaviartThomasNodal}, has been added. This feature is currently limited to meshes in standard orientation and affine geometries. Full support and performance optimizations will be provided in a future release.
\item Selected matrix-free algorithms can now exploit additional data locality
  between the matrix-vector product and vector operations happening nearby in
  an algorithm. Interfaces have been added to both the
  \texttt{PreconditionChebyshev} class (a frequently used smoother in multigrid
  methods) and the conjugate gradient implementation in \texttt{SolverCG}.
  From the user's perspective, an operator needs to define a \texttt{vmult} operation,
  taking two additional \texttt{std::function} arguments. The first function
  defines the operation to be scheduled on the vector entries before the
  matrix-vector product touches them, and the second what happens
  afterwards. The new features also include a renumbering to maximize data
  locality. The theory is described in the
  contribution~\cite{kronbichler2022cg}.
\end{itemize}
Besides these new features, we improved the performance of
hanging-node-constraint evaluation on the CPU. Instead of performing
quasi-dense matrix-vector multiplications~\cite{KronbichlerKormann2012}, we now use an
approach based on in-place interpolation and sum factorization, similar
to what was already done in the GPU code~\cite{ljungkvist2017matrix}. In \cite{munch2022hn}, the algorithm is described
and performance numbers are shown, indicating a reduction
of overhead of cells with hanging nodes by a factor of ten.

Finally, we have performed a major restructuring of internals
of the \texttt{FEEvaluation} classes. This reduces some overheads for low polynomial degrees and will enable us to add support for new element types in the future.

@@

\begin{c++}
fe\_eval.evaluate(false, true, true)          // old (deprecated)
fe\_eval.evaluate(EvaluationFlags::gradients) // new
\end{c++}

@@

In release 9.3~\cite{dealII93}, we added support for global-coarsening multigrid in
addition to the established local-smoothing infrastructure. Global
coarsening algorithms smoothen on the whole computational domain on each
multigrid level, which is obtained by coarsening the finest cells of
the next finer multigrid level.
For this purpose, we use a sequence of triangulations, and we perform
the smoothing only on their active levels. To create the sequence of
triangulations, one can use the functions \texttt{MGTransferGlobalCoarseningTools::create\_geometric\_coarsening\_sequence()}. A new version takes an
instance of \texttt{RepartitioningPolicyTools::Base} (see Subsection~\ref{sec:repartitioning}) as argument, which allows to specify the parallel
distribution of each multigrid level (in contrast to the fixed first-child policy in the case of local smoothing). These features have been developed and tuned for runs on supercomputer scale
with complicated coarse meshes as presented in \cite{kronbichler2021next}.
Furthermore, we added support for block vectors,
fixed a number of limitations, and performed performance optimizations of
the transfer operator; particularly, the redundant copy from/to temporary vectors
has been eliminated. Furthermore, hanging-node constraints are applied
efficiently in the same way as in the matrix-free loops (see Subsection~\ref{sec:mf}).

In \cite{munch2022gc}, the performance of the local-smoothing and global-coarsening
infrastructure of \texttt{deal.II} was compared for locally refined meshes. The results indicate that
the local definition of multigrid levels might introduce load imbalances
in the case of local smoothing so that global coarsening is favorable despite
potentially more expensive intergrid transfers. In order to judge the benefits
of one approach against the other, \texttt{deal.II} provides new functions
\texttt{workload\_imbalance()} and \texttt{vertical\_communication\_efficiency()}
in the \texttt{MGTools} namespace for the  estimation of the imbalance during, e.g.,
smoothing and the
communication efficiency during intergrid transfer purely based on the given mesh.
% MK: I would not add this part, it does not really fit into this paper as
% there is nothing to report at this point (and we should then add references
% to the actual literature).
%The publication \cite{munch2022gc} also points out that not
%all types of smoothers are applicable for global coarsening due to the
%presence of hanging nodes, which is a motivation to add new smoother types
%to \dealii in the future.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Integration of the Computational Geometry Algorithms
  Library (CGAL)}\label{sec:cgalwrappers}

The Computational Geometry Algorithms Library (CGAL, \url{https://www.cgal.org/}) is a widely used
library to describe geometries and meshes \cite{cgal}. \dealii{} now
has wrappers for CGAL classes and functions, provided in the new
namespace \texttt{CGALWrappers}, and implementing functionality
spanning from mesh generation to boolean operations between \dealii{}
triangulations and cells. These wrappers are enabled only if \dealii{}
is compiled with \texttt{C++17} and, for the first time in \dealii{}'s
25-year history, provides a built-in ability to generate meshes for
arbitrary geometries.

@@

// 5) convert surface to volume mesh via surface_mesh_to_volumetric_mesh()
  \end{c++}
  The output of the boolean operation can be seen in Fig.~\ref{fig:corefinement}, while in Fig.~\ref{fig:corefinement_remeshed}
  the same mesh has been remeshed.
  \begin{figure}
    \centering
    \begin{subfigure}[b]{0.28\textwidth}
      \centering
      \includegraphics[width=\textwidth]{png/heart_implicit.png}
      \caption{\label{fig:heart_tria}}
    \end{subfigure}\qquad
    \hfill
    \begin{subfigure}[b]{0.28\textwidth}
      \centering
      \includegraphics[width=\textwidth]{png/intersection_cube_sphere_mesh.png}
      \caption{\label{fig:corefinement}}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.35\textwidth}
      \centering
      \includegraphics[width=\textwidth]{png/cube_sphere_remeshed.png}
      \caption{ \label{fig:corefinement_remeshed}}
    \end{subfigure}
    \caption{\it (a) \texttt{Triangulation<3>} of a heart surface starting from an implicit function $f$. (b) Union of the unit cube with an hyper ball. Notice how the co-refinement around the intersection gives badly shaped elements. (c) The remeshed \dealii{} Triangulation.}
  \end{figure}

\texttt{CGALWrappers::compute\_quadrature\_on\_boolean\_operation()} returns a \texttt{Quadrature<3>} that allows exact integration on polyhedral elements coming out of a \texttt{BooleanOperation} between \dealii cells.
The quadrature rule is built by meshing the polyhedral region with tetrahedra, computing on each tetrahedron a \texttt{QGaussSimplex<3>} quadrature rule by using \texttt{QSimplex<3>::\allowbreak{}compute\_affine\_transformation()}, and finally
collecting all of the rules together, giving a \texttt{Quadrature<3>} formula on the \emph{physical} element.

These utilities will be the building blocks for adding functions to the \todo{NonMatching} namespace that will assemble coupling terms like $(u,v)_{B}$,\marginpar{The $B$ and $\Omega$ here need to be streamlined with \eqref{eq:boundingbox}, right now they are the opposite} with $B$ a domain immersed in a fixed background mesh $\Omega$ and $u,v$ finite element functions on $V_h(\Omega)$, as needed, e.g., in 
the context of CutFEM (see Section~\ref{sec:cut}) of Nitsche's method to weakly impose boundary conditions at an interface. The same applies to coupling terms of the form $(u,q)_B$ in formulations using Lagrange multipliers, where now $q \in Q_h(B)$, with $Q_h(B)$ the space of the multiplier variable.
Note that the most relevant difference between this and the \texttt{QuadratureGenerator} in Section~\ref{sec:cut} is that the \texttt{Quadrature} objects are created directly from two overlapping grids, one
spanning over $B$ and the other one over $\Omega$, and not from a level set function.

@@

In the release described herein, we have reorganized the data
structures and optimized the algorithms that support using particles
in \dealii{}. In our previously reported improvements~\citep{dealII93} we stored all particle data as a separate contiguous array for each particle property, but particle identifiers (IDs) were still stored in a multimap tree-like structure.

@@

In addition to the new storage structure, we have made the following algorithmic improvements:
Determining if a particle is inside a cell after changing its position involves inverting the mapping for this cell. We have reorganized our algorithms to perform these inversions on a batch of particles in the same cell instead of particle-by-particle, which allows us to make use of vectorized instructions during the inversion using the generic scheme of~\cite{KronbichlerKormann2012}.
In addition, after sorting all particles into their new cells, the arrays that store particle properties are now sorted in the same order as the particle IDs in the list of arrays, which allows for cache efficient iteration over particle properties. To avoid a costly sorting operation this operation is executed as a copy of the existing data into a new data container that replaces the existing container\todo{@RG: I don't get this sentence.}.

@@

In order to justify the work the developers of \dealii{} put into this
software, we ask that papers using the library reference one of the
\dealii{} papers. This helps us justify the effort we put into this library.

>>>>

\documentclass{ansarticle-preprint}
%\usepackage{ucs}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
%\usepackage{cite}
\usepackage{anslistings}
\usepackage{multicol}
\usepackage{pdfsync}
\usepackage{enumitem}

@@

\title{The \dealii Library, Version 9.4}

@@

   \author[4]{Marco~Feder}
\affil[4]{SISSA,
   International School for Advanced Studies,
   Via Bonomea 265,
   34136, Trieste, Italy.
   {\texttt{marco.feder/luca.heltai@sissa.it}}}

@@

\author[5]{Rene~Gassm{\"o}ller}
\affil[5]{Department of Geological Sciences,
   University of Florida,
   1843 Stadium Road,
   Gainesville, FL, 32611, USA.
  {\texttt{rene.gassmoeller@ufl.edu}}}

\author[6]{Timo~Heister}
 \affil[6]{School of Mathematical and Statistical Sciences,
   Clemson University,
   Clemson, SC, 29634, USA
   {\texttt{heister@clemson.edu}}}

@@

 \author[7,8]{Martin~Kronbichler}
 \affil[7]{Department of Information Technology,
   Uppsala University,
   Box 337, 751\,05 Uppsala, Sweden.
   {\texttt{martin.kronbichler/simon.sticko@it.uu.se}}}
 \affil[8]{Institute of Mathematics,
   University of Augsburg,
   Universit\"atsstr.~12a, 86159 Augsburg, Germany.
   {\texttt{martin.kronbichler@uni-a.de}}}

\author[9]{Matthias~Maier}
\affil[9]{Department of Mathematics,
  Texas A\&M University,
  3368 TAMU,
  College Station, TX 77845, USA.
  {\texttt{maier@math.tamu.edu}}}

\author[8,10]{Peter Munch}
 \affil[10]{Institute of Material Systems Modeling,
 Helmholtz-Zentrum Hereon,
 Max-Planck-Str. 1, 21502 Geesthacht, Germany.
   {\texttt{peter.muench@hereon.de}}}

\author[11]{Jean-Paul~Pelteret}
\affil[11]{Independent researcher.
{\texttt{jppelteret@gmail.com}}}

\author[7,12]{Simon~Sticko}
\affil[12]{Department of Mathematics and Mathematical Statistics,
   Umeå University,
   SE-90187 Umeå, Sweden}

@@

\author[13]{David Wells}
\affil[13]{Department of Mathematics, University of North Carolina,
  Chapel Hill, NC 27516, USA.
  {\texttt{drwells@email.unc.edu}}}

@@

\dealii version 9.4.0 was released June 10, 2022.
\todo{Update release date.}
This paper provides an
overview of the new features of this release and serves as a citable
reference for the \dealii software library version 9.4. \dealii is an
object-oriented finite element library used around the world in the
development of finite element solvers. It is available for free under the
GNU Lesser General Public License (LGPL). Downloads are available at
\url{https://www.dealii.org/} and \url{https://github.com/dealii/dealii}.

The major changes of this release are:
%
\begin{itemize}
  \item Advances in simplex- and mixed-mesh support (see Section~\ref{sec:simplex});
  \item Repartitioning of distributed meshes (see Section~\ref{sec:repartitioning});
  \item Advances in matrix-free infrastructure (see Section~\ref{sec:mf});
  \item Advances in multigrid infrastructure (see Section~\ref{sec:multigrid});
  \item CutFEM support (see Section~\ref{sec:cut});
  \item Experimental integration of the Computational Geometry Algorithms Library (CGAL) (see Section~\ref{sec:cgalwrappers});
  \item Performance improvements in the particle infrastructure (see Section~\ref{sec:particles});
  \item Improvements to unstructured communication (see Section~\ref{sec:CA});
  \item Three new tutorial programs and one new code gallery program (see Section~\ref{subsec:steps}).
\end{itemize}
%

While all of these major changes are discussed in detail in
Section~\ref{sec:major}, there
are a number of other noteworthy changes in the current \dealii release,
which we briefly outline in the remainder of this section:
%
\begin{itemize}
%  \item \texttt{AffineConstraints::make\_consistent\_in\_parallel()} allows
%  to make constraints consistent in parallel.
  \item The \texttt{DataOutResample} class interpolates values defined on one
  triangulation onto a second potentially unrelated triangulation.
  By using this class, one can output the result obtained on an
  unstructured mesh on a structured one (which might facilitate a more
  memory-efficient storage format, for example if this second
  triangulation is a uniformly refined rectangle or box), or one can create a slice in 3D.
  \item The new member function \texttt{find\_point\_owner\_rank()} of
        \texttt{parallel\allowbreak ::distributed::\allowbreak Tri\allowbreak angulation} allows one to find the MPI
  ranks of cells containing specified points.
  It is communication-free and leverages the functionality of \pfrst (>v.2.2).
  Its algorithm is described in \cite{burstedde2020parallel}. This information will
  enable efficient construction of the
  communication pattern used in the class \texttt{Utilties::\allowbreak MPI::\allowbreak RemotePointEvaluation}. Furthermore, this function could be used in the future to allow
  particle simulations in which particle movement is not
  limited by CFL conditions, as done in \cite{mirzadeh2016parallel}.
  \item The new function
  \texttt{GridGenerator::pipe\_junction()}
  generates a triangulation of three cone-shaped pipes that cross at a bifurcation point in any possible configuration.
  A manifold description is applied to the boundary, which can be extended into the volume via transfinite interpolation \cite{Gordon82} using the \texttt{TransfiniteInterpolationManifold} class \cite{dealII90}.
  \item A new DoF renumbering function \texttt{DoFRenumbering::support\_point\_wise()} which groups together
  shape functions by their support point. This functionality is useful in both developing nodal schemes since, e.g., the $x$, $y$, and $z$
  velocities at a point will be consecutive in the solution vector. It also improves interoperability with external libraries which expect
  data in this format.
  \item A new module \texttt{Utilities::MPI::LargeCount} which enables sending and receiving MPI messages containing more than $2^{31}$ objects.
  This library either uses the new MPI-4 functions, such as \texttt{MPI\_Send\_c()}, or an internal implementation of large objects for
  compatibility with MPI-3.
  \item The \texttt{FEInterfaceValues} class, which computes common quantities at the interface of two cells, has been overhauled to make it
  more consistent with the rest of the library and use more intuitive names for functions. For example,
  \texttt{FEInterfaceValues::jump\_gradient()} is now \texttt{FEInterfaceValues::jump\_in\_shape\_gradients()}. Several new
  functions, such as \texttt{FEInterfaceValues::get\_jump\_in\_function\_values()}, have also been added.
  \item Vectors attached to \texttt{DataOut} do not need to be in ghosted state anymore. Internally, we create a copy of the vector with appropriate ghosting.
\end{itemize}
%
The changelog lists more than 100 other features and bugfixes.

@@

This release of \dealii contains a number of large and significant changes,
which will be discussed in this section.
It of course also includes a
vast number of smaller changes and added functionality; the details of these
can be found
\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_3_0_and_9_4_0.html}
{in the file that lists all changes for this release}; see \cite{changes94}.

@@

\dealii has supported simplex and mixed meshes since the previous
release, 9.3, on an experimental basis.
We have continued to work on this support; the current state is
sufficient to support several larger-scale programs, although not all
functionality in \dealii works with such meshes yet. Specifically,
we have fixed many bugs and generalized existing functions that
previously only worked for hypercube-shaped cells. The most notable
new functionalities are:

@@

\dealii has two classes that support distributed storage of meshes:
\texttt{parallel::\allowbreak distributed::\allowbreak Triangulation}
(which in the following we will abbreviate as \texttt{p::d::T}) and
\texttt{parallel::\allowbreak fully\allowbreak distributed::\allowbreak
  Triangulation} (or, in short, \texttt{p::f::T}). The former is
partitioned based on the space-filling Morton (or Z-order) curve
implemented by the \pfrst backend. The latter, more recent
class is currently statically partitioned at creation time.

@@

The matrix-free infrastructure of \dealii is used to solve problems
without assembling matrices, providing only the action of the matrix
instead. It has seen a number of new features in this release, the most
notable ones being:
\begin{itemize}
\item Improved support for computations with Hessians of shape functions: just like values and gradients, Hessians can be
now evaluated and integrated during matrix-free loops
both for cells and faces. This could enable, for example, writing
a matrix-free version of step-47, which solves the biharmonic equation with the discontinuous Galerkin method.
\item Cell-centric loops now also allow access to gradients and Hessians
of neighboring cells on faces. The major difficulty here was the relative orientation
of cells on unstructured meshes.
\item Users can now create their own cell batches, by providing \texttt{FEEvaluation::reinit()} a list of cell IDs. \texttt{FEEvaluation}
accesses the appropriate data and reshuffles mapping data accordingly on
the fly in order to enable vectorization over cells. The new feature is useful in several
contexts. Examples are simulations with sharp interfaces (e.g., two-phase flow
or shock capturing), where one needs to treat cells that are ``cut'' by
the interface in a special way. A challenge is that cell batches
might contain cut or non-cut cells, making the vectorization of these operations potentially more complicated. Previous functionality has provided the option of masking certain cells in cell batches, which works well if
the code paths do not diverge too much. Another way is to
categorize cells during \texttt{MatrixFree::reinit()} in such a way that mixed cell batches do not occur. However, \texttt{MatrixFree::reinit()}
might be too expensive if recategorization needs to happen very frequently to follow the dynamics of a system, e.g., in each time step. Despite some overhead compared to static matrix-free loops, the new feature can be the best option in these scenarios.
\item Initial support for H(div)-conforming elements with Piola transform, based on Raviart--Thomas finite element spaces with the updated class \texttt{FE\_RaviartThomasNodal}, has been added. This feature is currently limited to meshes in standard orientation and affine geometries. Full support and performance optimizations will be provided in a future release.
\item Selected matrix-free algorithms can now exploit additional data locality
  between the matrix-vector product and vector operations happening nearby in
  an algorithm. Interfaces have been added to both the
  \texttt{PreconditionChebyshev} class (a frequently used smoother in multigrid
  methods) and the conjugate gradient implementation in \texttt{SolverCG}.
  From the user's perspective, an operator needs to define a \texttt{vmult} operation,
  taking two additional \texttt{std::function} arguments. The first function
  defines the operation to be scheduled on the vector entries before the
  matrix-vector product touches them, and the second what happens
  afterward. The new features also include a renumbering to maximize data
  locality. The theory is described in the
  contribution~\cite{kronbichler2022cg}.
\end{itemize}
Besides these new features, we improved the performance of
hanging-node-constraint evaluation on the CPU. Instead of performing
quasi-dense matrix-vector multiplications~\cite{KronbichlerKormann2012}, we now use an
approach based on in-place interpolation and sum factorization, similar
to what was already done in the GPU code~\cite{ljungkvist2017matrix}. In \cite{munch2022hn}, the algorithm is described
and performance numbers are shown, indicating a reduction
of overhead of cells with hanging nodes by a factor of ten.

Finally, we have performed a major restructuring of the internals
of the \texttt{FEEvaluation} classes. This reduces some overheads for low polynomial degrees and will enable us to add support for new element types in the future.

@@

\begin{c++}
fe_eval.evaluate(false, true, false)         // old (deprecated)
fe_eval.evaluate(EvaluationFlags::gradients) // new
\end{c++}

@@

In release 9.3~\cite{dealII93}, we added support for global-coarsening multigrid in
addition to the established local-smoothing infrastructure. Global
coarsening algorithms smoothen over the whole computational domain on each
multigrid level, which is obtained by coarsening the finest cells of
the next finer multigrid level.
For this purpose, we use a sequence of triangulations, and we perform
the smoothing only on their active levels. To create the sequence of
triangulations, one can use the functions \texttt{MGTransferGlobalCoarseningTools::create\_geometric\_coarsening\_sequence()}. A new version takes an
instance of \texttt{RepartitioningPolicyTools::Base} (see Subsection~\ref{sec:repartitioning}) as argument, which allows specifying the parallel
distribution of each multigrid level (in contrast to the fixed first-child policy in the case of local smoothing). These features have been developed and tuned for running on a supercomputer scale
with complicated coarse meshes as presented in \cite{kronbichler2021next}.
Furthermore, we added support for block vectors,
fixed a number of limitations, and performed performance optimizations of
the transfer operator; particularly, the redundant copy from/to temporary vectors
has been eliminated. Furthermore, hanging-node constraints are applied
efficiently in the same way as in the matrix-free loops (see Subsection~\ref{sec:mf}).

In \cite{munch2022gc}, the performance of the local-smoothing and global-coarsening
infrastructure of \dealii was compared for locally refined meshes. The results indicate that
the local definition of multigrid levels might introduce load imbalances
in the case of local smoothing so that global coarsening is favorable despite
potentially more expensive intergrid transfers. In order to judge the benefits
of one approach against the other, \dealii provides new functions
\texttt{workload\_imbalance()} and \texttt{vertical\_communication\_efficiency()}
in the \texttt{MGTools} namespace for the  estimation of the imbalance during, e.g.,
smoothing or the
communication efficiency during intergrid transfer, purely based on the given mesh.
% MK: I would not add this part, it does not really fit into this paper as
% there is nothing to report at this point (and we should then add references
% to the actual literature).
%The publication \cite{munch2022gc} also points out that not
%all types of smoothers are applicable for global coarsening due to the
%presence of hanging nodes, which is a motivation to add new smoother types
%to \dealii in the future.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Experimental integration of the Computational Geometry Algorithms
  Library (CGAL)}\label{sec:cgalwrappers}

The Computational Geometry Algorithms Library (CGAL, \url{https://www.cgal.org/}) is a widely used
library to describe geometries and meshes \cite{cgal}. \dealii now
has wrappers for CGAL classes and functions, provided in the new
namespace \texttt{CGALWrappers}, and implementing functionality
spanning from mesh generation to boolean operations between \dealii
triangulations and cells. These wrappers are enabled only if \dealii
is compiled with \texttt{C++17}. \textit{Note: This feature is still experimental and
interfaces might change during the next release cycle.}

@@

// 5) convert surface to volume mesh via surface_mesh_to_volumetric_mesh()
\end{c++}
The output of the boolean operation can be seen in Fig.~\ref{fig:corefinement}, while in Fig.~\ref{fig:corefinement_remeshed}
the same mesh has been remeshed.
\begin{figure}
  \centering
  \begin{subfigure}[b]{0.28\textwidth}
    \centering
    \includegraphics[width=\textwidth]{png/heart_implicit.png}
    \caption{\label{fig:heart_tria}}
  \end{subfigure}\qquad
  \hfill
  \begin{subfigure}[b]{0.28\textwidth}
    \centering
    \includegraphics[width=\textwidth]{png/intersection_cube_sphere_mesh.png}
    \caption{\label{fig:corefinement}}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.35\textwidth}
    \centering
    \includegraphics[width=\textwidth]{png/cube_sphere_remeshed.png}
    \caption{ \label{fig:corefinement_remeshed}}
  \end{subfigure}
  \caption{\it (a) Triangulation created by filling a heart-shaped surface implicitly described by a function $f$. (b) Union of a cube with a sphere with badly shaped cells at the intersection. (c) Remeshed version of the same triangulation.}
\end{figure}

\texttt{CGALWrappers::compute\_quadrature\_on\_boolean\_operation()} returns a \texttt{Quadrature<3>} that allows exact integration of polynomials on polyhedral elements coming out of a \texttt{BooleanOperation} between \dealii cells.
The quadrature rule is built by meshing the polyhedral region with tetrahedra, computing on each tetrahedron a \texttt{QGaussSimplex<3>} quadrature rule by using \texttt{QSimplex<3>::\allowbreak{}compute\_affine\_transformation()}, and finally
collecting all of the rules together, giving a \texttt{Quadrature<3>} formula on the \emph{physical} element.

These utility functions will be the building blocks for functions in the \texttt{NonMatching} namespace that will, e.g.,  assemble coupling terms like $(u,v)_{\Omega}$, with $\Omega$ a domain immersed in a fixed background mesh $B$ and $u,v$ finite element functions on $V_h(B)$, as needed, e.g., in 
the context of CutFEM (see Section~\ref{sec:cut}) or Nitsche's method to weakly impose boundary conditions at an interface. The same applies to coupling terms of the form $(u,q)_{\Omega}$ in formulations using Lagrange multipliers, where now $q \in Q_h(\Omega)$, with $Q_h(\Omega)$ the space of the multiplier variable.
Note that the most relevant difference between this and the \texttt{QuadratureGenerator} in Section~\ref{sec:cut} is that the \texttt{Quadrature} objects are created directly from two overlapping grids, one
spanning over $B$ and the other one over $\Omega$, and not from a level set function.

@@

In the release described herein, we have reorganized the data
structures and optimized the algorithms that support using particles
in \dealii. In our previously reported improvements~\citep{dealII93} we stored all particle data as a separate contiguous array for each particle property, but particle identifiers (IDs) were still stored in a multimap tree-like structure.

@@

In addition to the new storage structure, we have made the following algorithmic improvements:
Determining if a particle is inside a cell after changing its position involves inverting the mapping for this cell. We have reorganized our algorithms to perform these inversions on a batch of particles in the same cell instead of particle-by-particle, which allows us to make use of vectorized instructions during the inversion using the generic scheme of~\cite{KronbichlerKormann2012}.
In addition, after sorting all particles into their new cells, the arrays that store particle properties are now sorted in the same order as the particle IDs in the list of arrays, which allows for cache efficient iteration over particle properties.
Because the particle IDs are already sorted in the intended order at this time, we can reorder the particle properties as part of a copy operation into a new data container that replaces the existing container. This approach avoids a costly in-place sort of the particle properties.

@@

In order to justify the work the developers of \dealii put into this
software, we ask that papers using the library reference one of the
\dealii papers. This helps us justify the effort we put into this library.
